Carbonate chemistry at Ny-Ålesund

### 2022-08-25 awipev-CO2.Rmd
air-sea fluxes final

Several salinity sensors are on site, 1 inside the Ferrybox (SBE45) and
5 outside (ctd183, ctd181, ctd578, ctd964/964b and ctd103) that are
switched periodically. Salinity from FB is taken in consideration. If
data gaps are encountered, they are filled with bottom salinity data
from ctd (> 8 m). New column "sal_fb_mix" is created and is the one used in
ESSD paper.

-   1 SBE45 inside the Ferrybox (sal_fb)
-   5 CTD > 8 m depth (sal_insitu_ctd)

@@

### **Depth profile**
This is done later.

@@

### **Comparison salinity FerryBox vs Salinity CTD REMOS 10 m** 
```{r, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# REVIEW OK?
reg <- ggreg(
  xdata = d_all$sal_insitu_ctd_10m,
  ydata = d_all$sal_fb,
  type = "MA",
  plot = TRUE,
  title = "Comparison salinity",
  #subtitle = "",
  xlab = "Salinity CTD REMOS 10 m",
  ylab = "Salinity FerryBox",
  proplot = TRUE,
  ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
print(reg$plot)

@@

Several temperature sensors are on site, 1 inside the Ferrybox (SBE45),
5 outside (ctd183, ctd181, ctd578, ctd964/964b and ctd103) that are
switched periodically and 1 fixed at 11 m depth (sbe38). Temperature in
situ at 11 m depth is taken in consideration. If data gaps are
encountered, they are filled with bottom temperature data from ctd (> 8
m). New column "temp_insitu_11m_mix" is created. For the moment, no gaps are
encountered and temp_insitu_11m_mix = temp_insitu_11m.

@@

### **Temperature Ferrybox vs Temperature in situ 11 m** 
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Remove temp_fb outliers 
d_all <- d_all %>%
  dplyr::mutate(
    temp_fb = ifelse(
      datetime >= as.POSIXct("2017-03-22 08:00:00") &
      datetime <= as.POSIXct("2017-03-23 16:00:00"),
      NA,
      temp_fb
    ),
    temp_fb = ifelse(temp_fb > 15, NA, temp_fb)
  )

@@

### **Discrete total alkalinity vs salinity **

@@

#### **Is there a seasonal change in the AT vs S relationship**
```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%", fig.align="center"}
d_all_at_discrete <- d_all_at_discrete %>% 
  dplyr::select(sal_fb_mix, at) %>% 
  drop_na() %>% 
  dplyr::mutate(cutoff = ifelse(sal_fb_mix < 34.5, "below", "above"))
below <- d_all_at_discrete %>% 
  dplyr::filter(cutoff == "below")
reg_below <- ggreg(
  xdata = below$sal_fb_mix,
  ydata = below$at,
  type = "OLS",
  plot = TRUE,
  title = "Discrete total alkalinity vs salinity below 34.5",
  #subtitle = "",
  xlab = "Salinity",
  ylab = TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"),
  proplot = TRUE,
  #ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
above <- d_all_at_discrete %>% 
  dplyr::filter(cutoff == "above")
reg_above <- ggreg(
  xdata = above$sal_fb_mix,
  ydata = above$at,
  type = "OLS",
  plot = TRUE,
  title = "Discrete total alkalinity vs salinity above 34.5",
  #subtitle = "",
  xlab = "Salinity",
  ylab = TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"),
  proplot = TRUE,
  #ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
g <- reg_below$plot | reg_above$plot
g
```

@@

30 min warm up --> (3 min Zeroing + 15 min Flushing + 702 minutes
Measuring) --> (3 + 15 + 702)...

@@

The relationship between the measured (Contros) and calculated pCO2
(blue line) is relatively poor. It is far from the 1:1 relationship
(black dashed line). The slope is
`r signif(reg$fit$regression.results[2,3], 3)`) and its 95% confidence
interval does include 1; r2 is `r round(reg$fit$rsquare, digits=2)`  (`r N = reg$fit$n`)`.

@@

offset_ph <- ph_sf %>% 
  drop_na() %>% 
  ggplot() +
  geom_point(aes(x=datetime, y=diff_spec_sf), size=1, color="blue") +
  geom_hline(yintercept=0, size=0.3) +
  geom_hline(yintercept=-0.02, size=0.3, linetype = "dashed") +
  geom_hline(yintercept=0.02, size=0.3, linetype = "dashed") +
  labs(x = "Date", y=expression(paste(pH[T], " difference")), title = NULL) +
  Mytheme(size_labs = 9, face_font = "plain") + theme(aspect.ratio=1)  + 
  scale_colour_discrete(guide="none") + 
  coord_fixed()
ggsave(filename = "figures/offset_ph.png", plot = offset_ph, width = 25, height = 25, units = "cm")
```
Offset between calibrated seaFET data and pH spectro. The offset must be between -0.2 and 0.2 pH units. "differences between calibrated sensor pH values and values obtained from discrete, check samples can be expected to be as small as 0.02 pH units, and perhaps less." (McLaughlin et al. 2017).

@@

<!-- **Conclusion**: The range is significantly larger during some months -->
<!-- when the data deeper than 8 m are considered. Also, pH is sometimes -->
<!-- higher at the surface, at other times it is higher on the bottom (June). -->
<!-- No data are available at 1 m (actually 0 to 2 m). So, to compare surface -->
<!-- and bottom, I combine 1 and 3 m (0 to 4 m) and 9 m (below 8 m). -->

<!-- Depth and month of the year have a statistically significant effect on -->
<!-- pH. The interaction is also significant, indicating that the effect of -->
<!-- depth differs depending on the month. -->

<!-- -   To compare with pCO2/pCO2 measured in the Ferrybox seems therefore -->
<!--     better to stick with pH_10m, that is at depths deeper than 8 m. -->
<!-- -   pCO2 would have the same behaviour as pH. For air-sea CO2 fluxes, -->
<!--     one should use data from 0 to 4 m. -->

@@

```{r durafet clean up, echo=FALSE, message= FALSE, warning=FALSE}
d_all <- d_all %>% 
  dplyr::mutate(pH_dur = ph_dur_corr) %>% # the right durafet pH at in situ temp from now on
  dplyr::select(-ph_dur, -ph_dur_corr)
# normalized durafet pH 4°C
tmp <- d_all %>%
  dplyr::select(datetime, at, sal_fb_mix, temp_fb, pH_dur) %>%
  tidyr::drop_na()
tmp$pH_dur_temp_4 <- seacarb::pHinsi(pH=tmp$pH_dur, ALK=tmp$at, Tinsi=4, Tlab=tmp$temp_fb, Pinsi=11/10, S=tmp$sal_fb_mix, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")
d_all <- d_all %>% 
  left_join(tmp %>% dplyr::select(datetime, pH_dur_temp_4), by="datetime")
```
## Range per month

@@

### **seaFET vs durafet pHs at 4°C (all > 8 m)**

@@

### **Spectro vs durafet**
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / durafet
reg <- ggreg(
  xdata = d_all$ph_s_dur_temp_fb,
  ydata = d_all$pH_dur,
  type = "MA",
  plot = TRUE,
  title = "pH spectro vs pH durafet (both at FB temperature)",
  xlab = "pH durafet",
  ylab = "pH spectro",
  proplot = TRUE
)
print(reg$plot + xlim(c(7.8, 8.35)) + ylim(c(7.8, 8.35)))
```

### **Spectro vs seafet**
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / seafet
reg <- ggreg(
  xdata = d_all$ph_s_sf_temp_insi,
  ydata = d_all$pHint_tot_sf,
  type = "MA",
  plot = TRUE,
  title = "pH spec. vs pH seafet (at in situ temperature)",
  xlab = "Spectrophotometric pH",
  ylab = "SeaFET pH",
  proplot = TRUE
)
pHspec_pHseafet <- reg$plot + xlim(c(8.05, 8.20)) + ylim(c(8.05, 8.20))
pHspec_pHseafet
#ggsave(filename = "figures/pHspec_pHcalc.png", plot = pHspec_pHcalc)

pHspec_pHseafet <- d_all %>% 
  dplyr::select(ph_s_sf_temp_insi, pHint_tot_sf) %>% 
  drop_na() %>% 
  ggplot(aes(x = ph_s_sf_temp_insi, y = pHint_tot_sf)) +
  geom_point(col="blue", size = 0.5) +
  xlim(8.05, 8.18) + ylim(8.05, 8.18) +
  labs(x="Spectrophotometric pH",
       y= "SeaFET pH")  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label), digits = 4)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 9) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
ggsave(filename = "figures/pHspec_pHseafet.png", plot = pHspec_pHseafet, width = 25, height = 25, units = "cm")
```
The relationship between spectrophotometric pH and SeaFET pH) and 
(blue line) is relatively good The slope is
`r signif(reg$fit$regression.results[2,3], 3)`) and its 95% confidence
interval does include 1; r2 is `r round(reg$fit$rsquare, digits=2)`  (`r N = reg$fit$n`)`.

### **Spectro vs calculated AT-CT**
```{r, eval=FALSE, echo=FALSE, message= FALSE, warning=FALSE}
# do not run because only 2 data points available...
dplyr::select(d_all, datetime, ph_s_sf_temp_4) %>% drop_na()
dplyr::select(d_all, datetime, pH_calc_fb_temp_4) %>% dplyr::filter(datetime > as.POSIXct ("2017-08-23")) %>% drop_na()

@@

ggsave(filename = "figures/k1k2_all_pco2_at.png", plot = p, width = 25, height = 20, units = "cm")
```
`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis using all pCO2 data.", format = "simple")`
`r knitr::kable(diff_lp_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Papadimitriou using all pCO2 data.", format = "simple")`
`r knitr::kable(diff_lm_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Millero using all pCO2 data.", format = "simple")`
`r include_graphics("figures/k1k2_all_pco2_at.png")`

@@

## Stratification
### Stratification salinity
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at the distribution of data
depth_labels <- c("0-2", "2-4", "4-6", "6-8", ">8")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

@@

### Stratification temperature
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, parse = TRUE, label = "T['surface'] > T['deep']") +
  annotate("text", x = 2, y = -1.8, parse = TRUE, label = "T['surface'] <  T['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
```

### Stratification density
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at density at all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = pressure_insitu_ctd)) %>% 
  group_by(depth_group, montht)
p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$density ~ tmp$depth_group * tmp$montht))

@@

# Stratification plots -------------------
```{r, include=FALSE}
# main deltas
g <- sal_delta_plt / temp_delta_plt / den_delta_plt
g <- g + plot_annotation(tag_levels = 'a')
print(g)
ggsave(filename = "figures/deltas.png", plot = g, width = 15, height = 20, units = "cm")
# pCO2 deltas
g <-  pco2_delta_plt / pco2_4_delta_plt
g <- g + plot_annotation(tag_levels = 'a')
print(g)
ggsave(filename = "figures/deltas_pco2.png", plot = g, width = 15, height = 20, units = "cm")
```

@@

```{r Takahashi thermal vs non-thermal, include=FALSE}
# According to Takahashi et al. (2002) and McCutcheon et al. (2021)
# Select data
taka <- d_all %>% # just for this chunk
  dplyr::select(datetime, monthd, sal_fb_mix, temp_insitu_11m, pco2_fb_insi_temp) %>% 
  dplyr::mutate(year = lubridate::year(datetime),
                season = dplyr::case_when(monthd %in% c(12, 1, 2) ~ "winter",
                                          monthd %in% c(3, 4, 5) ~ "spring",
                                          monthd %in% c(6, 7, 8) ~ "summer",
                                          monthd %in% c(9, 10, 11) ~ "fall")) %>% 
  # eliminate years 2015 and 2021 which are incomplete
  dplyr::filter(datetime > as.POSIXct("2016-01-01 00:00:00") &
                  datetime < as.POSIXct("2021-01-01 00:00:00"))

@@

# Save all ----------------------------------------------------------------

>>>>

Several salinity sensors are on site, 1 inside the Ferrybox (SBE45) and
5 outside (ctd183, ctd181, ctd578, ctd964/964b and ctd103) that are
switched periodically. Salinity from FB is taken in consideration. If
data gaps are encountered, they are filled with bottom salinity data
from ctd (\> 8 m). New column "sal_fb_mix" is created and is the one
used in ESSD paper.

-   1 SBE45 inside the Ferrybox (sal_fb)
-   5 CTD \> 8 m depth (sal_insitu_ctd)

@@

### **Depth profile**

This is done later.

@@

### **Comparison salinity FerryBox vs Salinity CTD REMOS 10 m**

```{r, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# REVIEW OK?
reg <- ggreg(
  xdata = d_all$sal_insitu_ctd_10m,
  ydata = d_all$sal_fb,
  type = "MA",
  plot = TRUE,
  title = "Comparison salinity",
  #subtitle = "",
  xlab = "Salinity CTD REMOS 10 m",
  ylab = "Salinity FerryBox",
  proplot = TRUE,
  ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
print(reg$plot)

@@

Several temperature sensors are on site, 1 inside the Ferrybox (SBE45),
5 outside (ctd183, ctd181, ctd578, ctd964/964b and ctd103) that are
switched periodically and 1 fixed at 11 m depth (sbe38). Temperature in
situ at 11 m depth is taken in consideration. If data gaps are
encountered, they are filled with bottom temperature data from ctd (\> 8
m). New column "temp_insitu_11m_mix" is created. For the moment, no gaps
are encountered and temp_insitu_11m_mix = temp_insitu_11m.

@@

### **Temperature Ferrybox vs Temperature in situ 11 m**

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Remove temp_fb outliers 
d_all <- d_all %>%
  dplyr::mutate(
    temp_fb = ifelse(
      datetime >= as.POSIXct("2017-03-22 08:00:00") &
      datetime <= as.POSIXct("2017-03-23 16:00:00"),
      NA,
      temp_fb
    ),
    temp_fb = ifelse(temp_fb > 15, NA, temp_fb)
  )

@@

### **Discrete total alkalinity vs salinity**

@@

#### **Is there a seasonal change in the AT vs S relationship**

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, out.width="100%", fig.align="center"}
d_all_at_discrete <- d_all_at_discrete %>% 
  dplyr::select(sal_fb_mix, at) %>% 
  drop_na() %>% 
  dplyr::mutate(cutoff = ifelse(sal_fb_mix < 34.5, "below", "above"))
below <- d_all_at_discrete %>% 
  dplyr::filter(cutoff == "below")
reg_below <- ggreg(
  xdata = below$sal_fb_mix,
  ydata = below$at,
  type = "OLS",
  plot = TRUE,
  title = "Discrete total alkalinity vs salinity below 34.5",
  #subtitle = "",
  xlab = "Salinity",
  ylab = TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"),
  proplot = TRUE,
  #ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
above <- d_all_at_discrete %>% 
  dplyr::filter(cutoff == "above")
reg_above <- ggreg(
  xdata = above$sal_fb_mix,
  ydata = above$at,
  type = "OLS",
  plot = TRUE,
  title = "Discrete total alkalinity vs salinity above 34.5",
  #subtitle = "",
  xlab = "Salinity",
  ylab = TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"),
  proplot = TRUE,
  #ylim(c(30.5,35.5)) + xlim(c(28,35.5)),
  coord_fixed(ratio = 1)
)
g <- reg_below$plot | reg_above$plot
g
```

@@

30 min warm up --\> (3 min Zeroing + 15 min Flushing + 702 minutes
Measuring) --\> (3 + 15 + 702)...

@@

The relationship between the measured (Contros) and calculated pCO2
(blue line) is relatively poor. It is far from the 1:1 relationship
(black dashed line). The slope is
`r signif(reg$fit$regression.results[2,3], 3)`) and its 95% confidence
interval does include 1; r2 is `r round(reg$fit$rsquare, digits=2)`
(`r N = reg$fit$n`)\`.

@@

offset_ph <- ph_sf %>% 
  drop_na() %>% 
  ggplot() +
  geom_point(aes(x=datetime, y=diff_spec_sf), size=1, color="blue") +
  geom_hline(yintercept=0, size=0.3) +
  geom_hline(yintercept=-0.02, size=0.3, linetype = "dashed") +
  geom_hline(yintercept=0.02, size=0.3, linetype = "dashed") +
  labs(x = "Date", y=expression(paste(pH[T], " difference")), title = NULL) +
  Mytheme(size_labs = 9, face_font = "plain") + theme(aspect.ratio=1)  + 
  scale_colour_discrete(guide="none") + 
  coord_fixed()
ggsave(filename = "figures/offset_ph.png", plot = offset_ph, width = 25, height = 25, units = "cm")
```

Offset between calibrated seaFET data and pH spectro. The offset must be
between -0.2 and 0.2 pH units. "differences between calibrated sensor pH
values and values obtained from discrete, check samples can be expected
to be as small as 0.02 pH units, and perhaps less." (McLaughlin et al.
2017).

@@

<!-- **Conclusion**: The range is significantly larger during some months -->

<!-- when the data deeper than 8 m are considered. Also, pH is sometimes -->

<!-- higher at the surface, at other times it is higher on the bottom (June). -->

<!-- No data are available at 1 m (actually 0 to 2 m). So, to compare surface -->

<!-- and bottom, I combine 1 and 3 m (0 to 4 m) and 9 m (below 8 m). -->

<!-- Depth and month of the year have a statistically significant effect on -->

<!-- pH. The interaction is also significant, indicating that the effect of -->

<!-- depth differs depending on the month. -->

<!-- -   To compare with pCO2/pCO2 measured in the Ferrybox seems therefore -->

<!--     better to stick with pH_10m, that is at depths deeper than 8 m. -->

<!-- -   pCO2 would have the same behaviour as pH. For air-sea CO2 fluxes, -->

<!--     one should use data from 0 to 4 m. -->

@@

```{r durafet clean up, echo=FALSE, message= FALSE, warning=FALSE}
d_all <- d_all %>% 
  dplyr::mutate(pH_dur = ph_dur_corr) %>% # the right durafet pH at in situ temp from now on
  dplyr::select(-ph_dur, -ph_dur_corr)
# normalized durafet pH 4°C
tmp <- d_all %>%
  dplyr::select(datetime, at, sal_fb_mix, temp_fb, pH_dur) %>%
  tidyr::drop_na()
tmp$pH_dur_temp_4 <- seacarb::pHinsi(pH=tmp$pH_dur, ALK=tmp$at, Tinsi=4, Tlab=tmp$temp_fb, Pinsi=11/10, S=tmp$sal_fb_mix, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")
d_all <- d_all %>% 
  left_join(tmp %>% dplyr::select(datetime, pH_dur_temp_4), by="datetime")
```

## Range per month

@@

### **seaFET vs durafet pHs at 4°C (all \> 8 m)**

@@

### **Spectro vs durafet**

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / durafet
reg <- ggreg(
  xdata = d_all$ph_s_dur_temp_fb,
  ydata = d_all$pH_dur,
  type = "MA",
  plot = TRUE,
  title = "pH spectro vs pH durafet (both at FB temperature)",
  xlab = "pH durafet",
  ylab = "pH spectro",
  proplot = TRUE
)
print(reg$plot + xlim(c(7.8, 8.35)) + ylim(c(7.8, 8.35)))
```

### **Spectro vs seafet**

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / seafet
reg <- ggreg(
  xdata = d_all$ph_s_sf_temp_insi,
  ydata = d_all$pHint_tot_sf,
  type = "MA",
  plot = TRUE,
  title = "pH spec. vs pH seafet (at in situ temperature)",
  xlab = "Spectrophotometric pH",
  ylab = "SeaFET pH",
  proplot = TRUE
)
pHspec_pHseafet <- reg$plot + xlim(c(8.05, 8.20)) + ylim(c(8.05, 8.20))
pHspec_pHseafet
#ggsave(filename = "figures/pHspec_pHcalc.png", plot = pHspec_pHcalc)

pHspec_pHseafet <- d_all %>% 
  dplyr::select(ph_s_sf_temp_insi, pHint_tot_sf) %>% 
  drop_na() %>% 
  ggplot(aes(x = ph_s_sf_temp_insi, y = pHint_tot_sf)) +
  geom_point(col="blue", size = 0.5) +
  xlim(8.05, 8.18) + ylim(8.05, 8.18) +
  labs(x="Spectrophotometric pH",
       y= "SeaFET pH")  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  #ggpmisc::stat_ma_eq(method = "lmodel2:MA", aes(label = after_stat(eq.label), digits = 4)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  theme_bw() + Mytheme(size_labs = 9) + theme(aspect.ratio = 1) +
  coord_fixed(ratio = 1, expand = TRUE, clip = "on")
ggsave(filename = "figures/pHspec_pHseafet.png", plot = pHspec_pHseafet, width = 25, height = 25, units = "cm")
```

The relationship between spectrophotometric pH and SeaFET pH) and (blue
line) is relatively good The slope is
`r signif(reg$fit$regression.results[2,3], 3)`) and its 95% confidence
interval does include 1; r2 is `r round(reg$fit$rsquare, digits=2)`
(`r N = reg$fit$n`)\`.

### **Spectro vs calculated AT-CT**

```{r, eval=FALSE, echo=FALSE, message= FALSE, warning=FALSE}
# do not run because only 2 data points available...
dplyr::select(d_all, datetime, ph_s_sf_temp_4) %>% drop_na()
dplyr::select(d_all, datetime, pH_calc_fb_temp_4) %>% dplyr::filter(datetime > as.POSIXct ("2017-08-23")) %>% drop_na()

@@

ggsave(filename = "figures/k1k2_all_pco2_at.png", plot = p, width = 25, height = 20, units = "cm")
```

`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis using all pCO2 data.", format = "simple")`
`r knitr::kable(diff_lp_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Papadimitriou using all pCO2 data.", format = "simple")`
`r knitr::kable(diff_lm_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Millero using all pCO2 data.", format = "simple")`
`r include_graphics("figures/k1k2_all_pco2_at.png")`

@@

## Stratification

### Stratification salinity

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at the distribution of data
depth_labels <- c("0-2", "2-4", "4-6", "6-8", ">8")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

@@

### Stratification temperature

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, parse = TRUE, label = "T['surface'] > T['deep']") +
  annotate("text", x = 2, y = -1.8, parse = TRUE, label = "T['surface'] <  T['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
```

### Stratification density

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at density at all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = pressure_insitu_ctd)) %>% 
  group_by(depth_group, montht)
p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$density ~ tmp$depth_group * tmp$montht))

@@

---- Stratification plots -------------------

```{r, include=FALSE}
# main deltas
g <- sal_delta_plt / temp_delta_plt / den_delta_plt
g <- g + plot_annotation(tag_levels = 'a')
print(g)
ggsave(filename = "figures/deltas.png", plot = g, width = 15, height = 20, units = "cm")
# pCO2 deltas
g <-  pco2_delta_plt / pco2_4_delta_plt
g <- g + plot_annotation(tag_levels = 'a')
print(g)
ggsave(filename = "figures/deltas_pco2.png", plot = g, width = 15, height = 20, units = "cm")
```

@@

```{r, include=FALSE}
# According to Takahashi et al. (2002) and McCutcheon et al. (2021)
# Select data
taka <- d_all %>% # just for this chunk
  dplyr::select(datetime, monthd, sal_fb_mix, temp_insitu_11m, pco2_fb_insi_temp) %>% 
  dplyr::mutate(year = lubridate::year(datetime),
                season = dplyr::case_when(monthd %in% c(12, 1, 2) ~ "winter",
                                          monthd %in% c(3, 4, 5) ~ "spring",
                                          monthd %in% c(6, 7, 8) ~ "summer",
                                          monthd %in% c(9, 10, 11) ~ "fall")) %>% 
  # eliminate years 2015 and 2021 which are incomplete
  dplyr::filter(datetime > as.POSIXct("2016-01-01 00:00:00") &
                  datetime < as.POSIXct("2021-01-01 00:00:00"))

@@

Save all ----
